---
title: "data_adding_noises"
author: "Yinqiu He"
date: "7/5/2023"
output: html_document
---


This file contains codes how to read and preprocess the raw data. 

## 1. Exposure Variable

- FOLIOCC: subject index
- MEOHP: exposure variable of interest (log transformation applied)


```{r}
exposure_data <- read.csv(file="P01_Phthalate_T1_012020.csv")
exposure_data <-  exposure_data[c("FOLIOCC","MEOHP")] 

#Apply log transformation to the original exposure variable 
exposure_data[,2]=log(exposure_data[,2])
```


## 2. Outcome Variable

- FOLIOCC: subject index
- cbmi: bmi of subjects 

```{r}
outcome_data <- read.csv("bmi.csv")
outcome_data <-  outcome_data[c("FOLIOCC", "cbmi")]  #add age to adjust
```

## 3. Mediators Variable

```{r}
mediator_data <- read.csv(file="Metab_Lipids.csv")
```

## 4. Covariates to be adjusted

- FOLIOCC: subject index
- DELIVERY_SEX: gender
- age: age
- Intercept: all values are 1

```{r}
other_covariates_tmp <- read.csv("bmi.csv")
other_covariates <- cbind(other_covariates_tmp[c("FOLIOCC", "DELIVERY_SEX", "age")], 
                          rep(1,nrow(other_covariates_tmp)))
colnames(other_covariates)[4] <- "Intercept"
```


## 6. Add Gaussian Noises to Data
```{r}
d1 = dim(exposure_data)
exposure_data[,2] = exposure_data[,2] + rnorm( d1[1] ) 

d2 = dim(outcome_data)
outcome_data[,2] = outcome_data[,2] + rnorm(d2[1])

d3 = dim(mediator_data)
mediator_data[,2:d3[2]] = mediator_data[,2:d3[2]] + matrix(rnorm(d3[1]*(d3[2]-1)), d3[1], d3[2]-1)
  
d4 = dim(other_covariates)
other_covariates[,2:3] = other_covariates[,2:3] + matrix(rnorm(d4[1]*2), d4[1], 2)
```

## 6. Preview data

We show the first 3 rows of data in 1-4 parts above.
```{r}
head(exposure_data, 3)
head(outcome_data, 3)
head(mediator_data[,1:4], 3) #only first 4 columns are shown 
head(other_covariates, 3)
```




## 7. Save Data
```{r}
# saveRDS(exposure_data, file = "exposure_data.Rds")
# saveRDS(outcome_data, file = "outcome_data.Rds")
# saveRDS(mediator_data, file = "mediator_data.Rds")
# saveRDS(other_covariates, file = "other_covariates.Rds")
```


