---
title: "Codes for Data Analysis"
author: ""
date: ""
output: html_document
---

<br>

- We next present 6 parts of data analysis in the following order. 
     
     
       1. Read Data.
       2. Marginal Test Results.
       3. Screening based on marginal tests. 
       4. Joint Model and Global Test.
       5. Test Individual Mediation Effect under Joint Model.
       6. Sensitivity Analysis. 


- Supplementary script files mentioned below can be found under the folder  `5_codes_for_each_step`. 

<br>

====================================================

# 0. Preparation 

====================================================

#### Install and Load Packages 
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(Rcpp)
library(RcppArmadillo)
library(mediation)
library(DBmypackage6two)
```

<br>
<br>

====================================================

# I. Read Data

====================================================

- The data used below is processed as in `data_adding_noises.Rmd`, where standard Gaussian noises are added to raw data. 
- Results in the manuscript are obtained by the raw data, which is processed as in `data_preprocessing.Rmd`, where no noises are added. 


```{r}
exposure_data <- readRDS("3_processed_data/exposure_data.Rds")
outcome_data <- readRDS("3_processed_data/outcome_data.Rds")
mediator_data <- readRDS("3_processed_data/mediator_data.Rds")
other_covariates <- readRDS("3_processed_data/other_covariates.Rds")
```


###  Explanations of Variables:

- FOLIOCC: subject index
- MEOHP: exposure variable of interest (log transformation applied)
- cbmi: bmi of subjects 
- DELIVERY_SEX: gender
- age: age
- Intercept: all values are 1
- mediator_data: contains 149 lipids (mediator variables)


<br>
<br>


====================================================

# II. Marginal Test Results

====================================================

- See `II_marignal.R` for codes of conducting marginal screening tests. 

- Read output results. 

```{r}
res_max <- readRDS('4_numerical_results/realdata_max.Rds')
res_prod <- readRDS('4_numerical_results/realdata_prod.Rds')
res_other <- readRDS('4_numerical_results/realdata_other.Rds')
```

- In the following outputs, we use abbreviations for 8 tests. 
   - J-AB: Joint significance test with adaptive bootstrap
   - J-B: Joint significance test with classical nonparametric bootstrap
   - J-MaxP: Joint significance test using maximum p-value
   - P-AB: Product of coefficients test with adaptive bootstrap
   - P-B: Product of coefficients test with classical nonparametric bootstrap
   - P-Sobel: Sobel test
   - MT: Method in Huang (2019a)
   - CM: bootstrap method in the package `Causal Mediation Analysis`  Tingley et al. (2014)


```{r}
Num_mediator=149
num_test = 8
res_all=matrix(NA, Num_mediator, num_test)

#Estimated p-valuse for 149 mediators and 8 tests
for (j in 1:Num_mediator){
  #results of joint significance statistic
  res_all[j,1:3] = res_max[[j]]
  
  #results of product statistic
  res_all[j,4:6] = res_prod[[j]]
  
  #results of other tests
  res_all[j,7:8] = res_other[[j]]
}

colnametext = c("J-AB","J-B", "J-MaxP", "P-AB","P-B", "P-Sobel", "MT", "CMA")
colnames(res_all) = colnametext
```

<br>
<br>

=======================================================================

# III.  Screening based on marginal tests

=======================================================================

### Apply BH procedure to select lipids with FDR < 0.05

- Codes for conducting BH procedure. 

```{r}
#BH method
pBH_all = apply(res_all, 2, function(x){ p.adjust(x, method = "BH") } ) #calculate adjusted p-values by BH
sel_all <- apply(pBH_all, 2, function(x){ which(x<0.05) } )

#Find a union of lipids selected by candiate methods
all_sel_ind = c()
all_test_screen = c("J-AB", "J-MaxP", "P-AB","P-B", "P-Sobel", "CMA")
for(testname in all_test_screen){
  j = which(colnametext == testname)
  all_sel_ind = union(all_sel_ind, sel_all[[j]])
}

#all_sel_ind = c(3, 40, 41, 53, 132, 133, 136, 2, 14, 50, 52, 67, 74, 78, 95)
mediator_names <- colnames(mediator_data)[-1]
sel_med_names <- mediator_names[all_sel_ind] #names of the 15 selected lipids
```

- Combine selected lipids with corresponding p-values. 

<!-- The following codes show selected lipids along with p-values.``` -->

```{r}
sel_data <- NULL
for(testname in all_test_screen){
  j = which(colnametext == testname)
  sel_data <- cbind(sel_data, res_all[all_sel_ind, j],
                    1*(all_sel_ind%in%sel_all[[j]]))
}

sel_med_names_1 <- c("X16.0.LYSO.PC_3", "FA 7:0-OH._1", "FA 7:0-OH._2", "FA 18:3_1", 
                 "GLY", "GLY-H2O", "LAURIC.ACID", "X16.0.LYSO.PC_2",  "FA.16:0-OH", 
                 "LPC 16:1_3",  "LPC 18:2", "URSODIOL_2", "DECENOYL", "FA 12:0-DiC", "FA 18:0-DiC")
rownames(sel_data) <- sel_med_names_1
colnames(sel_data)<- rep(c("J-AB", "J-MaxP", "P-AB", "P-PB", "P-Sobel", "CMA"), each=2)
```

<!-- #### p-values of screened lipids in ```sel_data``` -->

- Output table with lipids and p-values. 

```{r}
knitr::kable(sel_data[,c(1,3,5,7,9,11)], digits=c(4,3,5,3,3,3), format.args = list(scientific = TRUE),  format = "simple")
```



<!-- - Each test has two columns. -->
<!-- - The first column shows estimated p-value for each lipid. -->
<!-- - The second column shows whether the test selects the lipid (1 for Yes, 0 for No) -->


<br>
<br>

============================================================

# IV. Joint Model and Global Test

============================================================

- Consider the following joint model of the 15 selected lipids: 
$$
\begin{aligned}
 &~ \mathrm{outcome} \sim \beta_1  M_1+\ldots + \beta_{15} M_{15} +   \mathrm{exposure} + X\\ 
&~ M_j \sim     \alpha_j \mathrm{exposure} + X, \quad \text{for }  j=1,\ldots, 15
\end{aligned}
$$
where $X$ represents other adjusted confounders including intercept, gender, and age, and $M_j$ represents the $j$-th mediator (lipid).

- Global test of the joint mediation effect: 
$$\sum_{j=1}^{15} \alpha_j \beta_j = 0.$$

- We provide codes in `IV_JointTest.R`. 



<br>
<br>

==================================================================================================

# V. Test Individual Mediation Effect under Joint Model

==================================================================================================

For $j=1,\ldots, 15$, test $$H_0: \alpha_j \beta_j = 0$$ separately.  

<br>

### V.i Read p-values

- See `V_JointIndividualTest.R` for codes to obtain p-values.  

```{r}
all_boot_pval <- readRDS("4_numerical_results/all_boot_pval.Rds") 

source('5_codes_for_each_step/pre_functions_mult.R')  
sel_med_names_1 <- c("X16.0.LYSO.PC_3", "FA 7:0-OH._1", "FA 7:0-OH._2", "FA 18:3_1", 
                 "GLY", "GLY-H2O", "LAURIC.ACID", "X16.0.LYSO.PC_2",  "FA.16:0-OH", 
                 "LPC 16:1_3",  "LPC 18:2", "URSODIOL_2", "DECENOYL", "FA 12:0-DiC", "FA 18:0-DiC")
rownames(all_boot_pval) <- sel_med_names_1
colnames(all_boot_pval) <- c("PoC-AB", "PoC-B", "PoC-Sobel", "JS-AB", "JS-B", "JS-MaxP", "MT", "CMA")
knitr::kable(all_boot_pval[,c("JS-AB", "JS-MaxP", "PoC-AB", "PoC-B", "PoC-Sobel",  "CMA")], digits=rep(4, 7), format = "simple", align=(c("r", rep("c", 6))), caption="P-Values for Testing Individual Mediation Effects.")
```


<br>

--- 

### V.ii Apply BH procedure

- We next present results after applying BH procedure with FDR level = 0.2.

```{r}
fdr_level <- 0.2
pBH_all = apply(all_boot_pval, 2, function(x){ p.adjust(x, method = "BH") } )
sel_all <- apply(pBH_all, 2, function(x){ which(x<fdr_level) } )

sel_ind_2 = c()
sel_ind_2 = union(sel_ind_2, sel_all$`PoC-AB`)
sel_ind_2 = union(sel_ind_2, sel_all$`PoC-B`)
sel_ind_2 = union(sel_ind_2, sel_all$`PoC-Sobel`)
sel_ind_2 = union(sel_ind_2, sel_all$`JS-AB`)
sel_ind_2 = union(sel_ind_2, sel_all$`JS-MaxP`)
sel_ind_2 = union(sel_ind_2, sel_all$`CMA`)

sel_res_2 <- cbind( 1*(sel_ind_2%in%sel_all$`PoC-AB`), 
                    1*(sel_ind_2%in%sel_all$`PoC-B`), 
                    1*(sel_ind_2%in%sel_all$`PoC-Sobel`),
                    1*(sel_ind_2%in%sel_all$`JS-AB`),
                    1*(sel_ind_2%in%sel_all$`JS-MaxP`),
                    1*(sel_ind_2%in%sel_all$`CMA`)
                    )
rownames(sel_res_2) <- rownames(all_boot_pval)[sel_ind_2]
colnames(sel_res_2) <- c("PoC-AB", "PoC-B", "PoC-Sobel", "JS-AB", "JS-MaxP", "CMA")
```

<br>

- Output selection results from BH procedure. 

- 1 indicates selection, 0 indicates no selection.

```{r}
sel_res_2
```

<br>
<br>

====================================================

# VI. Sensitivity Analysis

====================================================

- See `VI_sensitivity.R` for codes of conducting sensitivity analysis.


- Read results of sensitivity analysis.

```{r}
#read results of sensitivity analysis in `VI_sensitivity.R` 
#results are obtained when rho=0.01

res_sens <- readRDS('4_numerical_results/realdata_sens2.rds')

#read data
mediator_data <- readRDS("3_processed_data/mediator_data.Rds")
other_covariates <- readRDS("3_processed_data/other_covariates.Rds")


all_med_names <- colnames(mediator_data)
Num_mediator= 149

all_var_corr <- matrix(NA,Num_mediator,4)
alpha_val <- matrix(NA, Num_mediator, 2) #the first column is alpha.hat, and the second column is its p-value
beta_val <- matrix(NA, Num_mediator, 2) #the first column is beta.hat, and the second column is its p-value
med_val <- rep(NA, Num_mediator) #alpha.hat * beta.hat
```

<br>

- Calculate sample residual correlation by each mediator.

```{r}
for(j in 1:Num_mediator){
  data_model <- data_combine_function( j,  exposure_data, outcome_data, mediator_data, other_covariates)
  med.fit <- lm(mediator ~ exposure + age + SEX + 1, data = data_model)
  out.fit <- lm(outcome ~ mediator + exposure + age + SEX + 1, data = data_model)
  
  corval <- cor(med.fit$residuals, out.fit$residuals)
  var_out <- var(out.fit$residuals)
  var_med <- var(med.fit$residuals)
  
  all_var_corr[j,1] = corval
  all_var_corr[j,2] = var_out
  all_var_corr[j,3] = var_med
  
  med.fit.summary <- summary(med.fit)
  out.fit.summary <- summary(out.fit)
  alpha_val[j, ] <- med.fit.summary$coefficients[2, c(1,4)]
  beta_val[j, ] <- out.fit.summary$coefficients[2, c(1,4)]
  med_val[j] = alpha_val[j, 1] * beta_val[j, 1]
}
```

<br>

- Read $\rho_{min}$ obtained from sensitivity analysis. 

```{r}
min_rho_zero <- rep(NA, Num_mediator)
for(j in 1:Num_mediator){
  sens.out <- res_sens[[j]]$sens
  #keep the minimum rho that has confounded results
  min_rho_zero[j] = sens.out$err.cr.d
}

sens_res <- data.frame(sel_med_names, med_val[all_sel_ind],  alpha_val[all_sel_ind,], beta_val[all_sel_ind,], min_rho_zero[all_sel_ind],  all_var_corr[all_sel_ind,1])

greeks = c(alpha='\u03b1', tau='\u03c4', sigma='\u03c3', sigmaSq='\u03c3\u00B2', beta='\u03b2', gamma='\u03b3')

colnames(sens_res) = c("", "NIE", "alpha", "p_a", "beta", "p_b", "rho", "rho_m")

sens_res[,1] = c("X16.0.LYSO.PC_3", "FA 7:0-OH._1", "FA 7:0-OH._2", "FA 18:3_1", 
                 "GLY", "GLY-H2O", "LAURIC.ACID", "X16.0.LYSO.PC_2",  "FA.16:0-OH", 
                 "LPC 16:1_3",  "LPC 18:2", "URSODIOL_2", "DECENOYL", "FA 12:0-DiC", "FA 18:0-DiC")
```

<br>

- Output table of sensitivity analysis.

```{r}
colname_sel = c("", "ME", "$\\hat{\\alpha}$", "$p_{\\alpha}$", "$\\hat{\\beta}$", "$p_{\\beta}$", "$\\hat{\\rho}$", "$\\rho_{min}$")
knitr::kable(sens_res, digits=c(18,2,2,2,2,2,2), format = "simple", col.names = colname_sel)
```


<br>
<br>
<br>


